Discovery of s-process enhanced stars in the LAMOST survey

### 2018-02-23 article/ms.tex
fix latex

\author[Brodie.~J. Norfolk et al.]{Brodie.~J. Norfolk,$^{1}$\thanks{E-mail: bjee7@student.monash.edu (MU)}\thanks{This paper includes data gathered with the 6.5 meter Magellan Telescopes located at Las Campanas Observatory, Chile.}
Andrew.~R. Casey,$^{1,2}$
Matthew ~T. Miles,$^{1}$
Alex ~J. Kemp,$^{1}$ \newauthor
Amanda ~I. Karakas,$^{1}$,
Kevin ~C. Schlaufman$^{4}$,
Melissa Ness,$^{5}$,
Anna Y.~Q. Ho,$^{3}$, \newauthor
John ~C. Lattanzio,$^{1}$ and
Alexander ~P. Ji,$^{6}$
\\
$^{1}$School of Physics \& Astronomy, Monash University, Clayton 3800, Victoria, Australia\\
$^{2}$Faculty of Information Technology, Monash University, Clayton 3800, Victoria, Australia\\
$^{3}$Cahill Center for Astrophysics, California Institute of Technology, MC 249-17, 1200 E California Blvd, Pasadena, Ca, 91125, USA\\
$^{4}$Department of Physics and Astronomy, Johns Hopkins University, 3400 N Charles St., Baltimore, MD 21218, USA
\\
$^{5}$Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany
\\
$^{6}$Department of Physics, Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology,
Cambridge, MA 02139, USA
}
\date{Accepted XXX. Received YYY; in original form ZZZ}

@@

\section{Methods} \label{sec:methods}
\subsection{LAMOST analysis}
\subsubsection{Data-driven analysis}
The LAMOST (Large sky Area Multi-Object Fibre Spectrographic) survey released low-resolution ($\mathcal{R} \approx 1800$) optical spectra (3700\,\AA\ to 9000\,\AA) for over two million stars in their second data release \citep{luo2015}. Stellar parameters ($T_{\rm eff}$, $\log{g}$, [M/H], [$\alpha$/M]) were derived using a data-driven model constructed using high fidelity APOGEE labels of 9,952 giant stars in common between LAMOST and the APOGEE survey \citep{ho2017}. This data-driven model was used to estimate stellar labels for 454,180 giant stars. We note that this sample is exclusively giant stars, implying that we would not identify dwarf s-process enhanced stars. Cross-validation experience shows that the typical uncertainties are about 70\,K in effective temperature $\rm T_{\rm eff}$, 0.1\,dex in surface gravity $\log{g}$, 0.1\,dex in metallicity [M/H], and 0.04\,dex in the $\alpha$-element abundance relative to overall metallicity [$\alpha$/M]. These uncertainties are comparable to the APOGEE uncertainties \citep{alam2015}.

@@

@@

\begin{enumerate} 
\item Profile amplitude for both enhancement lines must be $A < -0.05$.
\item Both amplitudes must be measured within 3$\sigma$ ($|A|/\sigma _A$ < 3).
\item The wavelength at each absorption line must be within 2\,\AA\ ($\lambda$ < 2).
\item The reduced $\chi^2$ from \emph{The Cannon} must be $\chi_r^2 < 3$.
\item And, the LAMOST spectra must have a signal-to-noise ratio of $S/N > 30\,\textrm{pixel}^{-1}$.
\end{enumerate}
In addition, a visual inspection was undertaken to exclude any results containing false positives, candidates exhibiting data reduction issues, apparent absorption finer then the spectral resolution, and overly noisy normalised LAMOST spectra. This combined method generated 967 s-process rich candidates which are provided in Table 1 (online only material) including; 257 `barium stars' enhanced in both barium and strontium, 49 stars exhibiting exclusively barium enhancement, and 661 exhibiting exclusively strontium enhancement.

@@

%\begin{table*}
%\centering
%\caption{Properties of 967 s-process candidates, the table is available online in its entirety. Here we show a portion to demonstrate its style and content.}
%\label{table:table1}
%\begin{adjustbox}{width=1\textwidth}
%\begin{tabular}{@{}|l|l|c|c|c|c|c|c|c|c|c|c|c|c|c|@{}}
%\toprule
%2MASSID             & R.~A.         & Dec.        & $v_{r}$ & S/N & $\rm T_{\rm eff}$ & $\log{g}$ & [Fe/H] & [$\alpha$/Fe] & $\chi_r^2$ & [Ba/Fe] & [Sr/Fe] & \ion{Ba}{II} & \ion{Sr}{II} &  \ion{Ba}{II} \& \ion{Sr}{II} \\
%& (J2000) & (J2000) & (km\,s$^{-1}$) & (pixel$^{-1}$) & (K)  \\ \midrule
%J000019.26+501444.8 & 00:00:19.27 & +50:14:44.9 & -3.9  & 49      & 4973         & 3.27         & 0.21         & 0.08             & 0.66               & 0.25        & 0.83        & \ding{55} & \ding{51}  & \ding{55}   \\
%J000020.55+411348.1 & 00:00:20.56 & +41:13:48.2 & -28.2 & 32      & 4882         & 2.74         & -0.22        & 0.04             & 0.23               & -0.17       & 0.90        & \ding{55}& \ding{51}  & \ding{55}    \\
%J000134.95+490743.2 & 00:01:34.96 & +49:07:43.2 & -42.3 & 72      & 5044         & 3.11         & -0.54        & 0.11             & 0.79               & 1.02        & 0.45        & \ding{55} & \ding{55} & \ding{51}  \\
%J000258.09+410730.0 & 00:02:58.10 & +41:07:30.1 & -34.5 & 41      & 4697         & 2.57         & -0.22        & 0.12             & 0.98               & -0.10       & 0.80        & \ding{55} & \ding{51}  & \ding{55}  \\
%J000403.80+160257.1 & 00:04:03.80 & +16:02:57.2 & -35.4 & 42      & 5200         & 3.40         & -0.41        & 0.09             & 0.33               & 0.92        & 0.52        & \ding{55} & \ding{55} & \ding{51}  \\
%J000439.16+183350.3 & 00:04:39.17 & +18:33:50.4 & -38.7 & 31      & 4601         & 2.50         & 0.44         & 0.03             & 0.34               & -0.11       & 0.88        & \ding{55} & \ding{51}  & \ding{55}  \\
%J000444.87+400402.1 & 00:04:44.88 & +40:04:02.1 & -94.7 & 53      & 4172         & 1.46         & -0.08        & 0.09             & 0.72               & 0.03        & 0.91        & \ding{55} & \ding{51}  & \ding{55}  \\
%J000552.76+261849.3 & 00:05:52.76 & +26:18:49.4 & -22.5 & 73      & 4924         & 3.14         & -0.05        & 0.09             & 0.67               & 0.28        & 0.85        & \ding{55} & \ding{51}  & \ding{55}  \\
%J000737.70+394055.5 & 00:07:37.70 & +39:40:55.6 & -31.5 & 33      & 4700         & 2.80         & -0.04        & 0.07             & 0.28               & -0.13       & 0.81        & \ding{55} & \ding{51}  & \ding{55}  \\ \hline
%\end{tabular}
%\end{adjustbox}
%\end{table*}

@@

\subsection{Observations and data reduction}
On the night of 07 January 2018 we acquired a high-resolution spectrum of two barium star candidates (J09162834+0259348 and J08351472-0548480) using the MIKE (Magellan Inamori Kyocera Echelle) \citep{bernstein2003} spectrograph on the Magellan Clay telescope \citep{schectman2003} at Las Campanas Observatory, Chile. Candidates for follow-up observations were chosen based on observability constraints and ranked by their apparent visual magnitude, with both candidates having $V \approx 12$. We observed both stars in good seeing using the 0.7 arcsecond slit and 2x2 spatial on-chip binning, providing a spectral resolution of ($\mathcal{R} \approx 28,000$). Exposure times of 100 seconds were sufficient to achieve a S/N ratio exceeding 30 per pixel at 4500\,\AA. We acquired calibration (biases, milky, quartz, and Th-Ar arc lamp) frames in the afternoon.

We reduced the data using the \texttt{CarPy} Python package \citep{kelson2000,kelson2003}. We used spline functions to continuum-normalise individual echelle orders, and resampled the normalised spectra onto a uniform-sized wavelength map. We used a rest-frame normalised template of a FGK-type star to place the observed spectra in the rest frame.

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{hrd_new.pdf}
    \caption{Effective temperature $T_{\rm eff}$ and surface gravity $\log{g}$ for 257 candidate barium stars from LAMOST. Stars with $\log{g} > 2$ (as marked) have not evolved through the thermally pulsing asymptotic giant branch phase, and therefore their abundance signature must be the result of an extrinsic pollution event. The classification of stars with $\log{g} < 2$ would require additional data.}
    \label{fig:figure2}
\end{figure}

@@

\subsection{Carbon bands}
barium stars are excellent candidates for investigating the relationship between neutron-capture elements and other species that may be depleted or enhanced, since iron within these stars act as neutron seeds during the operation of the s-process. In the AGB phase the abundance of carbon is a product of helium fusion, specifically the triple-alpha process within a star. In an identical mechanism to the presence of s-process element enhancement, carbon enhanced stars exist intrinsically or extrinsically. Giants and supergiants become enhanced in carbon through TDU process discussed in Section \ref{sec:intro}, and extrinsic carbon enhancement occurs via mass transfer from a binary AGB star. Our analysis produced 178 CH and G band significantly enhanced stars (i.e., $[\textrm{C/Fe}] \gtrsim 0.5$) out of the total 257 barium candidates. These 178 stars exist in a metallicity range of 0 to -1.2, and can be termed "CH stars" are defined in previous literature \citep[e.g.][]{luck1991, mcclure1997}.
%Similar to the s-process enhancement, this suggests our candidates are consistent with being either intrinsic or extrinsic Barium stars.

\subsection{Comparison to AGB yields}
AGB stars, and chemically unusual stars that show the chemical signature of mass transfer pollution from binary companion AGB stars, highlight the variety of s-process enhancement. In Figure \ref{fig:figure3} we show the heavy-to-light s-process abundance ratio (taken as [Ba/Sr]) for all 257 barium stars identified in LAMOST. Although the [Ba/Sr] ratio is quite noisy, the overall metallicity [Fe/H] is quite precise (0.1\,dex). Despite our LAMOST sample of 454,180 giants having metallicities down to about $[{\rm Fe/H}] = -2$, we only detect barium stars down to metallicities of about $[{\rm Fe/H}] = -1.1$. Given our barium stars exist in this metallicity range of 0 to -1.1, we can dfeine them as CH stars as done in previous literature \citep[e.g.][]{luck1991, mcclure1997}.

The final surface s-process index [Ba/Sr], illustrated in Figure \ref{fig:figure3}, highlights the relationship between [Ba/Sr] yields and the mass of AGB stars. The Salpeter initial mass function (or any bottom-heavy initial mass function) predicts the majority of stars in the Milky Way will have lower masses, and as predicted the [Ba/Sr] values of our barium stars approximately fit the lower mass yields as seen in Figure \ref{fig:figure3}. However, it's important to note, our results do not indicate when the mass transfer occured, and our candidates fitting these theoratical yeilds may not have the correlating binary companion masses. Instead, mass transfer may have occured early on in the AGB S binary companion star's lifespan, and the resulting [Ba/Sr] abundance present in our barium candidates may relfect pollution from a lower ABG S star mass. Additionally, this can be extended to our candidates not approximated by theoratical yeilds. Other barium candidates not fitting this relationship represent the possibly more massive, and expectedly rarer, AGB S stars within our sample.

\begin{figure}
	\includegraphics[width=\columnwidth]{yields_test.pdf}
    \caption{Metallicities ([Fe/H]; x-axis) and heavy-to-light s-process abundance ratios ([Ba/Sr] as measured from LAMOST; y-axis) for 257 barium star candidates. Coloured lines blue, orange, and green, indicate surface [hs/ls] yields from \citet{cristallo2015} for different masses. Coloured lines red, purple, and brown indicate surface [hs/ls] yields from \citet{karakas_lugaro2016} for [Fe/H] at +0.3. 0.0, and -0.3. From \citet{karakas2018} for [Fe/H] at -0.7, and from \citet{fishlock2014} for [Fe/H] at -1.2.}
    \label{fig:figure3}
\end{figure}

@@

\section*{Acknowledgements}
We thank  Alex Ji (Carnegie Observatories), David W. Hogg (NYU), Melissa Ness (Columbia), and Hans-Walter Rix (MPIA) for useful discussions. 
A.~R.~C. is supported through an Australian Research Council Discovery Project under grant DP160100637.
A. Y. Q. H. was supported by the GROWTH project funded by the National Science Foundation under PIRE Grant No 1545949, and a National Science Foundation Graduate Research Fellowship under Grant No. DGE-1144469. 
This research has made use of NASA's Astrophysics Data System.
Guoshoujing Telescope (the Large Sky Area Multi-Object Fiber Spectroscopic Telescope LAMOST) is a National Major Scientific Project built by the Chinese Academy of Sciences. Funding for the project has been provided by the National Development and Reform Commission. LAMOST is operated and managed by the National Astronomical Observatories, Chinese Academy of Sciences. 
This work has made use of data from the European Space Agency (ESA) mission Gaia (http://www.cosmos.esa.int/gaia), processed by the Gaia Data Processing and Analysis Consortium (DPAC, http://www.cosmos.esa.int/web/gaia/dpac/consortium). Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the Gaia Multilateral Agreement.

@@

\bibitem[\protect\citeauthoryear{Kelson}{2003}]{kelson2003}
Kelson, D.~D. 2003, 
PASP, 115, 688

@@

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

>>>>

\author[Brodie.~J. Norfolk et al.]{Brodie.~J. Norfolk,$^{1}$\thanks{E-mail: bjee7@student.monash.edu (MU)}\thanks{This paper includes data gathered with the 6.5 meter Magellan Telescopes located at Las Campanas Observatory, Chile.}
Andrew.~R. Casey$^{1,2}$,
Matthew ~T. Miles$^{1}$,
Alex ~J. Kemp$^{1}$, \newauthor
Amanda ~I. Karakas$^{1}$,
Kevin ~C. Schlaufman$^{4}$,
Melissa Ness$^{5}$,
Anna Y.~Q. Ho$^{3}$, \newauthor
John ~C. Lattanzio$^{1}$, 
Alexander ~P. Ji$^{6}$
\\
$^{1}$School of Physics \& Astronomy, Monash University, Clayton 3800, Victoria, Australia\\
$^{2}$Faculty of Information Technology, Monash University, Clayton 3800, Victoria, Australia\\
$^{3}$Cahill Center for Astrophysics, California Institute of Technology, MC 249-17, 1200 E California Blvd, Pasadena, Ca, 91125, USA\\
$^{4}$Department of Physics and Astronomy, Johns Hopkins University, 3400 N Charles St., Baltimore, MD 21218, USA
\\
$^{5}$Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany
\\
$^{6}$Carnegie Observatories
}
\date{Accepted XXX. Received YYY; in original form ZZZ}

@@

\section{Methods} \label{sec:methods}
\subsection{LAMOST analysis}
\subsubsection{Data-driven analysis}
The LAMOST (Large sky Area Multi-Object Fibre Spectrographic) survey released low-resolution ($\mathcal{R} \approx 1800$) optical spectra (3700\,\AA\ to 9000\,\AA) for over two million stars in their second data release \citep{luo2015}. Stellar parameters ($T_{\rm eff}$, $\log{g}$, [M/H], [$\alpha$/M]) were derived using a data-driven model constructed using high fidelity APOGEE labels of 9,952 giant stars in common between LAMOST and the APOGEE survey \citep{ho2017}. This data-driven model was used to estimate stellar labels for 454,180 giant stars. We note that this sample is exclusively giant stars, implying that we would not identify dwarf s-process enhanced stars. Cross-validation tests shows that the typical uncertainties are about 70\,K in effective temperature $\rm T_{\rm eff}$, 0.1\,dex in surface gravity $\log{g}$, 0.1\,dex in metallicity [M/H], and 0.04\,dex in the $\alpha$-element abundance relative to overall metallicity [$\alpha$/M]. These uncertainties are comparable to the APOGEE uncertainties \citep{alam2015}.

@@

% FOR SOME REASON LATEX FALLS OVER IF WE DONT HAVE IT HERE
\begin{table}\end{table}

@@

\begin{enumerate} 
\item Profile amplitude for both enhancement lines must be $A < -0.05$.
\item Both amplitudes must be measured within 3$\sigma$ ($|A|/\sigma _A$ < 3).
\item The wavelength at each absorption line must be within 2\,\AA\ ($\lambda$ < 2).
\item The reduced $\chi^2$ from \emph{The Cannon} must be $\chi_r^2 < 3$.
\item And, the LAMOST spectra must have a signal-to-noise ratio of $S/N > 30\,\textrm{pixel}^{-1}$.
\end{enumerate}
In addition, a visual inspection was undertaken to exclude any results containing false positives, candidates exhibiting data reduction issues, apparent absorption finer then the spectral resolution, and overly noisy normalised LAMOST spectra. This combined method generated 967 s-process rich candidates which are provided in Table 1 (online only material) including; 257 `barium stars' enhanced in both barium and strontium, 49 stars exhibiting exclusively barium enhancement, and 661 with exclusively strontium enhancement.

@@

@@

\subsection{Observations and data reduction}
On the night of 07 January 2018 we acquired a high-resolution spectrum of two barium star candidates (J09162834+0259348 and J08351472-0548480) using the MIKE (Magellan Inamori Kyocera Echelle) \citep{bernstein2003} spectrograph on the Magellan Clay telescope \citep{schectman2003} at Las Campanas Observatory, Chile. Candidates for follow-up observations were chosen based on observability constraints and ranked by their apparent visual magnitude, with both candidates having $V \approx 12$. We observed both stars in good seeing using the 0.7 arcsecond slit and 2x2 spatial on-chip binning, providing a spectral resolution of ($\mathcal{R} \approx 28,000$). Exposure times of 100 seconds were sufficient to achieve a S/N ratio exceeding 30 per pixel at 4500\,\AA. We acquired calibration (biases, milky, quartz, and Th-Ar arc lamp) frames in the afternoon. We reduced the data using the \texttt{CarPy} package \citep{kelson2000}. We used spline functions to continuum-normalise all echelle orders, and resampled the normalised spectra onto a uniform-sized wavelength map. We used a rest-frame normalised template of a FGK-type star to place the observed spectra at rest.

@@

\begin{figure}
\centering
	\includegraphics[width=0.4\textwidth]{hrd_new.pdf}
    \caption{Effective temperature $T_{\rm eff}$ and surface gravity $\log{g}$ for 257 candidate barium stars from LAMOST.}
    %Stars with $\log{g} > 2$ (as marked) have not evolved through the thermally pulsing asymptotic giant branch phase, and therefore their abundance signature must be the result of an extrinsic pollution event. The classification of stars with $\log{g} < 2$ would require additional data.}
    \label{fig:figure2}
\end{figure}

@@

\begin{figure}
	\includegraphics[width=0.5\textwidth]{yields_test.pdf}
    \caption{Metallicities ([Fe/H]; x-axis) and heavy-to-light s-process abundance ratios ([Ba/Sr] as measured from LAMOST; y-axis) for 257 barium star candidates. Coloured lines blue, orange, and green, indicate surface [hs/ls] yields from \citet{cristallo2015} for different masses. Coloured lines red, purple, and brown indicate surface [hs/ls] yields from \citet{karakas_lugaro2016} for [Fe/H] at +0.3. 0.0, and -0.3. From \citet{karakas2018} for [Fe/H] at -0.7, and from \citet{fishlock2014} for [Fe/H] at -1.2.}
    \label{fig:figure3}
\end{figure}

\subsection{Carbon bands}
Barium stars are excellent candidates for investigating the relationship between neutron-capture elements and other species that may be depleted or enhanced, since iron within these stars act as neutron seeds during the operation of the s-process. In the AGB phase the abundance of carbon is a product of helium fusion, specifically the triple-alpha process within a star. In an identical mechanism to the presence of s-process element enhancement, carbon enhanced stars exist intrinsically or extrinsically. Giants and supergiants become enhanced in carbon through TDU process discussed in Section \ref{sec:intro}, and extrinsic carbon enhancement occurs via mass transfer from a binary AGB star. Our analysis produced 178 CH and G band significantly enhanced stars (i.e., $[\textrm{C/Fe}] \gtrsim 0.5$) out of the total 257 barium candidates. These 178 stars exist in a metallicity range of 0 to -1.2, and can be termed "CH stars" are defined in previous literature \citep[e.g.][]{luck1991, mcclure1997}.
%Similar to the s-process enhancement, this suggests our candidates are consistent with being either intrinsic or extrinsic Barium stars.

\subsection{Comparison to AGB yields}
AGB stars, and chemically unusual stars that show the chemical signature of mass transfer pollution from binary companion AGB stars, highlight the variety of s-process enhancement. In Figure \ref{fig:figure3} we show the heavy-to-light s-process abundance ratio (taken as [Ba/Sr]) for all 257 barium stars identified in LAMOST. Although the [Ba/Sr] ratio is quite noisy, the overall metallicity [Fe/H] is quite precise (0.1\,dex). Despite our LAMOST sample of 454,180 giants having metallicities down to about $[{\rm Fe/H}] = -2$, we only detect barium stars down to metallicities of about $[{\rm Fe/H}] = -1.1$. Given our barium stars exist in this metallicity range of 0 to -1.1, we can define them as CH stars as done in previous literature \citep[e.g.][]{luck1991, mcclure1997}.

The final surface s-process index [Ba/Sr], illustrated in Figure \ref{fig:figure3}, highlights the relationship between [Ba/Sr] yields and the mass of AGB stars. The Salpeter initial mass function (or any bottom-heavy initial mass function) predicts the majority of stars in the Milky Way will have lower masses, and as predicted the [Ba/Sr] values of our barium stars approximately fit the lower mass yields as seen in Figure \ref{fig:figure3}. However, it's important to note, our results do not indicate when the mass transfer occurred, and our candidates fitting these theoretical yields may not have the correlating binary companion masses. Instead, mass transfer may have occurred early on in the AGB S binary companion star's lifespan, and the resulting [Ba/Sr] abundance present in our barium candidates may reflect pollution from a lower ABG S star mass. Additionally, this can be extended to our candidates not approximated by theoretical yields. Other barium candidates not fitting this relationship represent the possibly more massive, and expectedly rarer, AGB S stars within our sample.

@@

\section*{Acknowledgements}
We thank David W. Hogg (NYU) and Hans-Walter Rix (MPIA) for useful discussions. 
A.~R.~C. is supported through an Australian Research Council Discovery Project under grant DP160100637.
A. Y. Q. H. was supported by the GROWTH project funded by the National Science Foundation under PIRE Grant No 1545949, and a National Science Foundation Graduate Research Fellowship under Grant No. DGE-1144469. 
This research has made use of NASA's Astrophysics Data System.
Guoshoujing Telescope (the Large Sky Area Multi-Object Fiber Spectroscopic Telescope LAMOST) is a National Major Scientific Project built by the Chinese Academy of Sciences. Funding for the project has been provided by the National Development and Reform Commission. LAMOST is operated and managed by the National Astronomical Observatories, Chinese Academy of Sciences. 
%This work has made use of data from the European Space Agency (ESA) mission Gaia (http://www.cosmos.esa.int/gaia), processed by the Gaia Data Processing and Analysis Consortium (DPAC, http://www.cosmos.esa.int/web/gaia/dpac/consortium). Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the Gaia Multilateral Agreement.

@@

%\bibitem[\protect\citeauthoryear{Kelson}{2003}]{kelson2003}
%Kelson, D.~D. 2003, 
%PASP, 115, 688

@@

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}
